---
title: "Fat Bear 2020-2021"
author: "Laura Mora"
author: "Maya Crandall-Malcolm"
author: "Mariem Snoussi"
date: "`r Sys.Date()`"
output: openintro::lab_report
              toc: true
            theme: united
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(mosaic)
library(modeest)
```


```{r}
bears<-read.csv("FullBearData.csv")
survey20<-read.csv("Survey2020.csv")
survey21<-read.csv("Survey2021.csv")
```

```{r}
# 2020 Bear 909 results
nineonine<-filter(survey20, BearNumber=="909")
mlv(nineonine$FurColor, method="mfv")
mlv(nineonine$SizeBefore, method="mfv")
mlv(nineonine$SizeAfter, method="mfv")
mean(nineonine$Angle)

# 2020 Bear 856 results
b856<-filter(survey20, BearNumber=="856")
mlv(b856$FurColor, method="mfv")
mlv(b856$SizeBefore, method="mfv")
mlv(b856$SizeAfter, method="mfv")
mean(b856$Angle)

#2020 bear 812
b812<-filter(survey20, BearNumber=="812")
mlv(b812$FurColor, method="mfv")
mlv(b812$SizeBefore, method="mfv")
mlv(b812$SizeAfter, method="mfv")
mean(b812$Angle)

# 2020 bear 747
b747<-filter(survey20, BearNumber=="747")
mlv(b747$FurColor, method="mfv")
mlv(b747$SizeBefore, method="mfv")
mlv(b747$SizeAfter, method="mfv")
mean(b747$Angle)

# 2020 bear 719
b719<-filter(survey20, BearNumber=="719")
mlv(b719$FurColor, method="mfv")
mlv(b719$SizeBefore, method="mfv")
mlv(b719$SizeAfter, method="mfv")
mean(b719$Angle)

# 2020 bear 480
b480<-filter(survey20, BearNumber=="480")
mlv(b480$FurColor, method="mfv")
mlv(b480$SizeBefore, method="mfv")
mlv(b480$SizeAfter, method="mfv")
mean(b480$Angle)

# 2020 bear 435
b435<-filter(survey20, BearNumber=="435")
mlv(b435$FurColor, method="mfv")
mlv(b435$SizeBefore, method="mfv")
mlv(b435$SizeAfter, method="mfv")
mean(b435$Angle)

# 2020 bear 402
b402<-filter(survey20, BearNumber=="402")
mlv(b402$FurColor, method="mfv")
mlv(b402$SizeBefore, method="mfv")
mlv(b402$SizeAfter, method="mfv")
mean(b402$Angle)

# 2020 bear 151
b151<-filter(survey20, BearNumber=="151")
mlv(b151$FurColor, method="mfv")
mlv(b151$SizeBefore, method="mfv")
mlv(b151$SizeAfter, method="mfv")
mean(b151$Angle)

# 2020 bear 128
b128<-filter(survey20, BearNumber=="128")
mlv(b128$FurColor, method="mfv")
mlv(b128$SizeBefore, method="mfv")
mlv(b128$SizeAfter, method="mfv")
mean(b128$Angle)

# 2020 bear 32
b32<-filter(survey20, BearNumber=="32")
mlv(b32$FurColor, method="mfv")
mlv(b32$SizeBefore, method="mfv")
mlv(b32$SizeAfter, method="mfv")
mean(b32$Angle)
```

```{r}
### 2021 survey results
# 812
b812<-filter(survey21, BearNumber=="812")
mlv(b812$FurColor, method="mfv")
mlv(b812$SizeBefore, method="mfv")
mlv(b812$SizeAfter, method="mfv")
mean(b812$Angle)

# 747
b747<-filter(survey21, BearNumber=="747")
mlv(b747$FurColor, method="mfv")
mlv(b747$SizeBefore, method="mfv")
mlv(b747$SizeAfter, method="mfv")
mean(b747$Angle)

# 634
b634<-filter(survey21, BearNumber=="634")
mlv(b634$FurColor, method="mfv")
mlv(b634$SizeBefore, method="mfv")
mlv(b634$SizeAfter, method="mfv")
mean(b634$Angle)

# 503
b503<-filter(survey21, BearNumber=="503")
mlv(b503$FurColor, method="mfv")
mlv(b503$SizeBefore, method="mfv")
mlv(b503$SizeAfter, method="mfv")
mean(b503$Angle)

# 480
b480<-filter(survey21, BearNumber=="480")
mlv(b480$FurColor, method="mfv")
mlv(b480$SizeBefore, method="mfv")
mlv(b480$SizeAfter, method="mfv")
mean(b480$Angle)

# 435 
b435<-filter(survey21, BearNumber=="435")
mlv(b435$FurColor, method="mfv")
mlv(b435$SizeBefore, method="mfv")
mlv(b435$SizeAfter, method="mfv")
mean(b435$Angle)

# 402
b402<-filter(survey21, BearNumber=="402")
mlv(b402$FurColor, method="mfv")
mlv(b402$SizeBefore, method="mfv")
mlv(b402$SizeAfter, method="mfv")
mean(b402$Angle)

# 151
b151<-filter(survey21, BearNumber=="151")
mlv(b151$FurColor, method="mfv")
mlv(b151$SizeBefore, method="mfv")
mlv(b151$SizeAfter, method="mfv")
mean(b151$Angle)

# 132
b132<-filter(survey21, BearNumber=="132")
mlv(b132$FurColor, method="mfv")
mlv(b132$SizeBefore, method="mfv")
mlv(b132$SizeAfter, method="mfv")
mean(b132$Angle)

# 131
b131<-filter(survey21, BearNumber=="131")
mlv(b131$FurColor, method="mfv")
mlv(b131$SizeBefore, method="mfv")
mlv(b131$SizeAfter, method="mfv")
mean(b131$Angle)

# 32
b32<-filter(survey21, BearNumber=="32")
mlv(b32$FurColor, method="mfv")
mlv(b32$SizeBefore, method="mfv")
mlv(b32$SizeAfter, method="mfv")
mean(b32$Angle)

# 128
b128<-filter(survey21, BearNumber=="128")
mlv(b128$FurColor, method="mfv")
mlv(b128$SizeBefore, method="mfv")
mlv(b128$SizeAfter, method="mfv")
mean(b128$Angle)
```


## Purpose
The purpose of this analysis is to evaluate what factors play a major role in the winner of Fat Bear Week. Fat Bear Week is an online single elimination tournament held in Katmai National Park, Alaska. This year it was held from September 29th to October 5th. According to the website where the voting takes place, who to vote for as the winner is subjective. Some voters may chose to vote based on dramatic change in weight over the year. Other voters might chose to vote based on additional challenges the bears may face in order to gain weight. But in general the idea is to vote for the fattest bear, in the voter's opinion. Variables that will be It is predicted that the more angled away from the camera the bear is, or the closer the butt of the bear is to the camera, the higher the likelihood of success. It is also predicted that the probability that a bear will win the second round increases as the number of votes that bear obtains in the first round increases. This analysis will also explore other variables that might play a role in the who named winner of Fat Bear Week.
.
## Data
.
# Population

# Outcome Variable
.
# Explanatory Variable


# Exploratory Visualizations
.

